Learning from Data Through Models"  # nolint: Grammarly.

### 2023-09-04 Paper/draft.Rmd
experimental design

An agent is of type $\theta \in \{\theta_L, \theta_M, \theta_H\}$  with $\theta_H > \theta_M > \theta_L$ and they face an unknown 
exogenous state
$\omega \in \{\omega_L, \omega_M, \omega_H\}$ with $\omega_H>\omega_M>\omega_L$. 
Each of the values of $\omega$ is realized with equal probability. The agent knows the distribution of $\omega$ but not its value. 
They also hold some prior belief about $\theta$ which is potentially misspecified.\footnote{The particular types of misspecification
that are allowed are discussed below.}
The agent chooses a binary gamble $e in \{e_L, e_M, e_H\}$ and observes whether the gamble is a success or a failure. '
If it is a success,
the agent gets a payoff of $1$ and if it is a failure, the agent gets a payoff of $0$. The probability of success is increasing in
both $\theta$ and $\omega$ and is fully described in the following table:

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & 30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & 20 \\
  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} 
\begin{tabular}{ c|c|c|c|}

Conditional on a type, the agent's flow payoff is maximized by choosing the gamble that 
matches the state. For example, if the value of $\omega$ is $\omega_H$, the agent's flow payoff is maximized by 
choosing $e_H$ and if the state is $\omega_L$ the flow payoff is maximized by choosing gamble $e_L$. The agent myopically chooses
gambles every period to maximize the flow payoff for $T<\infty$ periods, and the agent's payoff is the sum of the payoffs from 
each period.

After observing the outcome of each gamble, the agent updates their beliefs of both theta and omega in one of 4 different ways: 
As a \emph{dogmatic} as in @Heidhues2018 ; as a \emph{switcher} as in @Ba2023 ; as a \emph{self-attribution} updater as in @Coutts2020; 
or as a fully Bayesian agent. Each of these updating rules is described in detail below.

Notice that both $\theta$ and $\omega$ can be identified from the outcomes if enough variation in the effort choices exists.
This can be seen by confirming that no two columns are the same. Therefore, the agent can learn both their type and the 
state by observing enough outcomes of the gambles for different effort choices.

@@

$$\frac{p[h^t|\tilde{\theta}]}{p[h^t|\hat{\theta}]} > \alpha$$

@@

@@

>>>>

@@

$$\frac{p[h^t|\tilde{\theta}}{p[h^t|\hat{\theta}]} > \alpha$$

@@

# A Simple Example
The agent can be of one of 3 types: $\theta \in \{\theta_L, \theta_M, \theta_H\}$  with $\theta_H > \theta_M > \theta_L$.
They face an unknown exogenous success rate
$\omega \in \{\omega_L, \omega_M, \omega_H\}$ with $\omega_H>\omega_M>\omega_L$. Each of the values of $\omega$ is realized with equal probability. 
The agent knows the distribution of $\omega$ but not its realized value. Let the true type be $\theta^*$ 
and the true state be $\omega^*$.

The agent holds some prior belief about $\theta$ \footnote{which is potentially misspecified as in the dogmatic and switcher cases discussed} 
and chooses a binary gamble $e in \{e_L, e_M, e_H\}$. The agent observes whether the gamble is a success or a failure and 
gets a payoff of $1$ and if it is a success. They get $0$ otherwise.

The probability of success is increasing in
both $\theta$ and $\omega$ and is fully described in the following table:

  \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\omega_H$} & \multicolumn{1}{c}{$\omega_M$} & \multicolumn{1}{c}{$\omega_L$}\\
  \cline{2-4}
  $e_H$ & 50 & 20 & 2 \\
  \cline{2-4}
  $e_M$ & 45 & 30 & 7 \\
  \cline{2-4}
  $e_L$ & 40 & 25 & 20 \\

  \cline{2-4}
  \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\theta_L$} & \multicolumn{1}{c}{}\\
\end{tabular}
\hspace{.3cm} 
\begin{tabular}{ c|c|c|c|}

Conditional on a type, the agent's flow payoff is maximized by choosing the gamble that 
matches the state. For example, if the value of $\omega$ is $\omega_H$, the agent's flow payoff is maximized by 
choosing $e_H$ and if the state is $\omega_L$ the flow payoff is maximized by choosing gamble $e_L$. The agent myopically chooses
gambles every period to maximize the flow payoff for $T$ periods.

After observing the outcome of each gamble, the agent updates their beliefs using some algorithm and moves on to the next period.

Notice that both $\theta$ and $\omega$ can be identified from the outcomes if enough variation in the effort choices exists.
This can be seen by confirming there is no pair of $\theta$ and $\omega$ such that the probability of success is the same for all effort 
choices. Therefore, by changing the effort choice, the agent can learn both their type and the state if they observe enough outcomes.

In this example, for an agent with a dogmatic belief about their type, a self-defeating equilibrium is one in which the agent 
chooses an effort level that, under the true $\theta$, yields a frequency of success that is consistent with the agent's 
misspecified belief. That is, the observed rate of success under the true type and the effort level chosen, 
$P[\text{sucess}|\theta^*, e]$ is the same as the probability of success under the agent's belief, $P[\text{sucess}|\hat{\theta}, e]$.

In the data-generating process described above, there are six such equilibria. For example, if the agent is of type $\theta_M$  but
mistakenly believes that he is of type $\hat{\theta}=\theta_H$ and the 
and $\omega^* = \omega_M$, when the effort chosen is $e_L$, the agent will observe a success with 45% chance. Because the agent 
dogmatically believes that their type is high, they will erroneously conclude that the rate is $\omega_L$. Under this belief,
the optimal action is $e_L$ which will continue to generate successes with $45%$ probability, further reinforcing
the incorrect belief. If the agent were to learn the true type and rate, they would instead choose $e_M$, yielding a 
success with $65%$ probability and strictly higher expected payoffs.

From @Heidhues2018, we know that if $\theta^*$ and $\omega^*$ are such that there is a self-confirming equilibrium, a dogmatic agent will always
fall into the trap and continue to make sub-optimal choices. On the other hand, a switcher will be able to learn the true type and state
as long as their prior is diffused enough and their switching threshold is low enough as in @Ba2023.

By including self-confirming equilibria, the example captures the forces from each of the updating mechanisms discussed in the 
previous section and allows for direct comparison of all the theories. For realizations of $(\theta, \omega)$ for which there are 
self-confirming equilibria, the dogmatic agent will fall into the 
trap whereas the switcher will be able to escape it. Similarly, an agent with self-attribution bias will update their beliefs differently from an unbiased Bayesian, leading
them to choose different gambles. I exploit such cases in order to test which model is a better fit for how subjects behave in a 
laboratory experiment.

@@

I recruited XXX undergraduate subjects from the CESS lab at NYU who participated in an in-person experiment. Sessions lasted approximately 
XXX hours and subjects earned an average payment of XXX. The experiment was programmed using oTree [@chen2016otree].

The experiment consisted of 2 treatments: the ego-relevant condition and the stereotype condition. Subjects participated in only one
of them. Treatments were randomly assigned at the session level. The tasks were identical across treatments, but in the ego condition
the probability of success was based on the subject's own performance in a quiz, while in the stereotype condition, the probability
od success was based on the performance of a randomly selected subject from another session.

The experiment had 3 parts. In Part 1 subjects had 2 minutes to answer as many multiple-choice questions as they could from a 
20 question quiz in a particular topic. They did this for 6 different topics, one after the other. The topics were: Math, Verbal 
Reasoning, Pop-culture and Art, Science and Technology, US Geography, and Sports and Video Games. The maximum number of questions 
they could answer correctly was 20. However, they did not know how many questions were available and they were given no feedback.

After taking all 6 quizzes, they proceeded to part 2 where they were asked to guess their score on each of them. In the stereotype 
treatment they were additionally asked to guess the score of another participant. All they knew about the other participant was their
gender identity and whether they were US nationals or not.
For each guess they had three score options: 5 or fewer, between 6 and 15, 
16 or more. Each of these correspond to $\theta_L$, $\theta_M$, and $\theta_H$ respectively. 
They were also asked to say how confident they felt about their choices with 4 options: it was a random guess, there is another equally likely score, I am pretty sure, I am completely sure. These 4 answers are mapped to priors that place probabilities 
.33, .50, .75, and 1 to the chosen type. The remaining probability is split equally among the other two types. Questions in part 2
were not incentivized, but subjects were told that providing an accurate answer would increase their chances of earning more money in 
the last part of the experiment.

In part 3 subjects completed a belief updating task for each of the quizzes. Before starting the task they were reminded of their guess
for the score. In the ego treatment they were reminded of their guess about themselves and in the stereotype treatment they were reminded
of their guess about the other participant. In the stereotype treatment, they were also reminded of the characteristics of the other participant.

For one topic at a time and in random order,
they were presented with the three gambles from the example and were asked to choose one of them. The probability of success was 
determined by their own score in the ego-relevant condition and by the score of the other participant in the stereotype condition. 
Subjects had access to the three probability
tables in the printout on the instructions at all times and the meaning of each cell was explained in detail in the instructions. However, 
on the screen they had to choose which of the 3 tables they wanted to see before entering their choice in it. This was done as 
an alternative to a belief elicitation in each round. I take their choice of table to be a noisy measure of their beliefs about
the underlying type.

Once they have entered their choice, they observe a sample of 10 outcomes from the gamble they chose. After observing the outcomes, they returned to the choice screen and entered a new choice. In the choice screen 
subjects had access to the entire history of gambles and outcomes for that task. Once they entered 11 gambles (and observed 110 outcomes), they moved on to 
the next topic and repeated the same procedure. They all did so for all 6 topics.

At the end of the experiment, one of the 6 topics was randomly selected to determine the payment. They earned $\$.20$ for each 
correct answer in the quiz and for each success in part 3.
